Evolution of Processes in Delivering Reliable Software in the European Scientific Arena

### 2018-09-14 paper-ieee-trans.tex

\begin{figure*}[ht]
\centering
\includegraphics[width=\textwidth]{images/fig2.png}
\caption{Comparison of the unit testing coverage values for the {\sl INDIGO--DataCloud}
software stack over the {\sl INDIGO--1} and {\sl INDIGO--2} major releases. Products that
were exclusively released during {\sl INDIGO--2} do not show data for the first period of
{\sl INDIGO--1}.}
\label{fig:fig_unittest}
\end{figure*}

@@

\item \textit{Unit and Functional testing}.
Changes involving the addition of new functionalities are required to be tested. \textit{Regression
testing} in this context is accomplished by enforcing the definition and periodic execution
of the tests, covering past fixed issues, to ensure that they are not reintroduced during the development
activities. \textit{Unit testing} completes the source code testing evaluation by focusing on the code's
internal design. The SQA criteria of {\sl INDIGO--DataCloud} set a recommended threshold of 70\% coverage
for unit testing. In Fig.~\ref{fig:fig_unittest}, unit testing coverage is compared for each software
component between the two major releases, denoting an incremental trend over the course of the project:
code coverage increased from an average value of 50.32\% to 63.34\%, outlined by the respective dates of
the project's first (INDIGO--1) and second (INDIGO--2) releases. Nevertheless, Fig.~\ref{fig:fig_unittest}
shows a decrease in the unit testing coverage values for several products. As unit tests cover low-level
code elements, they are best considered while the new code is being written \cite{unit-test-frameworks}.
We have observed that this decrease was aligned with high-demanding periods of software development, as
seen in the previous months of the INDIGO-2 major release. Regardless of the potentially higher defect
density that this fact brought along, the overall performance kept progressing as the rest of the software
stack substantially improved their own unit testing coverage statistics. By INDIGO--2, 53\% of the software
components (17 out of 32) were over the threshold of the aforementioned project's SQA code coverage
recommendation (70\%), while 75\% (24 out of 32) of the product stack exceeded 50\% coverage. 
The individual records followed also a growing trend (only 6 out of 32 components had lower values in 
INDIGO--2). It is worth noting that Fig.~\ref{fig:fig_unittest} not only includes new software developments 
implemented from scratch within the project. Software tools and libraries from 
external open--source projects -- such as {\sl tosca-parser} or {\sl heat-translator} --, contributed upstream
by the project, and well--established products, involved in {\sl INDIGO--DataCloud}
but not contributing with the 100\% of their codebases -- such as {\sl dCache} --, are also considered in the
analysis. However the values given apply to the entire codebase, even for the latter case of products. These
products challenged the application of the SQA policies described in this section, by two means:
\begin{itemize}
    \item Previously developed code was not refactored since it was not \textit{owned} by the project.
    \item An agreement on prevailing SQA policies for such cases where quality practices were already in place.
\end{itemize}

@@

\bibitem{unit-test-frameworks}
Paul Hamill,
``Unit test frameworks: tools for high-quality software development,''
\emph{O'Reilly Media, Inc.},
2004.

>>>>

\begin{figure*}[ht]
\centering
\includegraphics[width=\textwidth]{images/fig2.png}
\caption{Comparison of the unit testing coverage values for the {\sl INDIGO--DataCloud}
software stack over the {\sl INDIGO--1} and {\sl INDIGO--2} major releases.}
\label{fig:fig_unittest}
\end{figure*}

@@

\item \textit{Unit and Functional testing}.
Changes involving the addition of new functionalities are required to be tested. \textit{Regression
testing} in this context is accomplished by enforcing the definition and periodic execution
of the tests, covering past fixed issues, to ensure that they are not reintroduced during the development
activities. \textit{Unit testing} completes the source code testing evaluation by focusing on the code's
internal design. The SQA criteria of {\sl INDIGO--DataCloud} set a recommended threshold of 70\% coverage
for unit testing. In Fig.~\ref{fig:fig_unittest}, unit testing coverage is compared for each software
component between the two major releases, denoting an incremental trend over the course of the project:
code coverage increased from an average value of 50.32\% to 63.34\%, outlined by the respective dates of
the project's first (INDIGO--1) and second (INDIGO--2) releases. By INDIGO--2, 53\% of the software components
(17 out of 32) were over the threshold of the aforementioned project's SQA code coverage recommendation
(70\%), while 75\% (24 out of 32) of the product stack exceeded 50\% coverage.
The individual records followed also a growing trend (only 6 out of 32 components had lower values in
INDIGO--2). It is worth noting that Fig.~\ref{fig:fig_unittest} not only includes new software developments
implemented from scratch within the project. Software tools and libraries from
external open--source projects -- such as {\sl tosca-parser} or {\sl heat-translator} --, contributed upstream
by the project, and well--established products, involved in {\sl INDIGO--DataCloud}
but not contributing with the 100\% of their codebases -- such as {\sl dCache} --, are also considered in the
analysis. However the values given apply to the entire codebase, even for the latter case of products. These
products challenged the application of the SQA policies described in this section, by two means:
\begin{itemize}
    \item Previously developed code was not refactored since it was not \textit{owned} by the project.
    \item An agreement on prevailing SQA policies for such cases where quality practices were already in place.
\end{itemize}

@@
